\documentclass[a4paper,12pt]{scrartcl}

\usepackage{bm,amsmath,url,graphicx}
\usepackage{palatino}
\usepackage{color, xcolor}
\usepackage{listings}


\newcommand{\n}{{\bf n}}
\newcommand{\h}{{\bf h}}
\newcommand{\x}{{\bf x}}
\newcommand{\w}{{\bf w}}
\newcommand{\HH}{{\bf H}}
\newcommand{\C}{{\bf C}}
\newcommand{\thb}{{\boldsymbol{\theta}}}
\newcommand{\mub}{{\boldsymbol{\mu}}}
\newcommand{\python}{{\fbox{\texttt{\bfseries python}}\quad}}
\newcommand{\pen}{{\fbox{\texttt{\bfseries pen\&paper}}\quad}}

\renewcommand{\familydefault}{\rmdefault}


\begin{document}
\section*{\bf SGN-41007 Pattern Recognition and Machine Learning}
\emph{Exercise Set 4: February 1--February 3, 2017}
\bigskip
\sloppy

\lstdefinestyle{mystyle}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=single,
  xleftmargin=\parindent,
  language=Python,
  showstringspaces=false,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
  moredelim=**[is][\color{red}]{@}{@},
}

\lstset{language=Python,style=mystyle} 


\noindent
Exercises consist of both pen\&paper and computer assignments.
Pen\&paper questions are solved at home before exercises, while
computer assignments are solved during exercise hours. The
computer assignments are marked by text \python and 
Pen\&paper questions by text \pen

\begin{enumerate}

\item \pen \emph{Design an LDA classifier manually.}

A dataset consists of two classes, whose distributions are
assumed Gaussian, and whose sample covariances and means are
the following:
\begin{alignat*}{2}
\mub_1 &= \begin{pmatrix}
-1\\
1
\end{pmatrix}
 && \mub_2= \begin{pmatrix}
-6\\
2
\end{pmatrix}\\
\C_1 &= \begin{pmatrix}
2 & 0.1 \\
0.1 & 0.2
\end{pmatrix}
\qquad 
&&\C_2 = \begin{pmatrix}
3 & -2 \\
-2 & 2
\end{pmatrix}
\end{alignat*}
Calculate the projection vector $\w$.
In order to be fully manual, invert the $2\times 2$ matrix using
the rule
\[
\begin{pmatrix}
	a & b\\
	c & d
\end{pmatrix}^{-1}
= 
\frac{1}{ad - bc}
\begin{pmatrix}
	d & -b \\
	-c & a
\end{pmatrix}
\]

\item \pen \emph{Compute the threshold and classify.}

The LDA decision rule requires a threshold $T$:
\begin{itemize}
	\item Decide class = 1 if $\w^T\x \ge T$.
	\item Decide class = 0 if $\w^T\x < T$.
\end{itemize}
Compute $T$ by setting it at the center of projected class means,
$\mub_1$ and $\mub_2$.

 Which class will be predicted for sample $\x = (1,2)$?

\item \pen \emph{Compute the threshold more properly and classify.}

The previous approach to defining the threshold $T$ did not take into account
the fact that the two classes have different spreads, and the threshold
should probably not be exactly at the center.

A more appropriate approach would thus compute the projection of
the multivariate Gaussians and set the threshold accordingly.
The projected Gaussians are univariate normal: ${\cal N}(\w^T\mub_1, \w^T\C_1\w)$
and ${\cal N}(\w^T\mub_2, \w^T\C_2\w)$. Formulate the classification 
problem as a likelihood ratio test and choose the threshold based on that.

Which class will be predicted for sample $\x = (1,2)$?

\item \python \emph{Extract Local Binary Pattern features for classification.}

In this exercise we will extract image features for categorization of traffic signs.
Download the following file:


\url{http://www.cs.tut.fi/courses/SGN-41007/GTSRB_subset.zip}

If has two folders each containing 100 images from the German Traffic Sign
Recognition Benchmark (GTSRB); a competition organized in IJCNN-2011 conference.

Load all images from the two folders and compute their Local Binary Pattern
features using \verb+skimage.feature.local_binary_pattern+.\footnote{See
\url{http://scikit-image.org/docs/dev/auto_examples/plot_local_binary_pattern.html}}
The function returns an image with same size as the original, so you will
also have to compute the histogram with \verb+numpy.histogram+.
Note that this is a similar task to exercise 4 last week.
The result should be a feature matrix \verb+X+ and label vector \verb+y+.

\item \python \emph{Train classifiers for the GTSRB task.}

Create a list of three classifiers with their default parameters:
\begin{itemize}
	\item \verb+sklearn.neighbors.KNeighborsClassifier+
	\item \verb+sklearn.lda.LDA+
	\item \verb+sklearn.svm.SVC+
\end{itemize}

Split your data into two parts---80\% for training and 20\% for testing---using
\verb+sklearn.cross_validation.train_test_split+. 

Train each classifier in a for loop and assess the accuracy in the test set
using \verb+sklearn.metrics.accuracy_score+.
Which one is the best?
\end{enumerate}

\end{document}          
